%\documentclass[paper]{geophysics}
\documentclass[manuscript,revised]{geophysics}

% An example of defining macros
\newcommand{\rs}[1]{\mathstrut\mbox{\scriptsize\rm #1}}
\newcommand{\rr}[1]{\mbox{\rm #1}}

\usepackage{amsfonts} 
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsmath}

\begin{document}

\title{Cross-correlation misfit measurement and adjoint source for full-waveform inversion}

\renewcommand{\thefootnote}{\fnsymbol{footnote}} 

\ms{GEO-Example} % manuscript number

\address{

\footnotemark[1]
Energy Geosciences Division,
Lawrence Berkeley National Laboratory,
One Cyclotron Road, Berkeley, CA 94720, USA.\\
E-mail: kechen@lbl.gov\\}
\author{Ke Chen\footnotemark[1]}

%\footer{Example}
\lefthead{}
\righthead{\emph  Cross-correlation misfit}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%ABSTRACT%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
The waveform least-squares misfit function is often used in full-waveform inversion. However, it may lead to local minimum if the observed data and synthetic data are cycle-skipped. This document shows the cross-correlation travel time misfit function and the corresponding adjoint source for synthetic and real seismic data. 
\end{abstract}


%%%%%%%%%%%%%%%%%THEORY%%%%%%%%%%%%%%%
\section{Theory}
\subsection{Cross-correlation}

The cross-correlation of observed and synthetic signal is defined as
\begin{equation}
\begin{aligned}
C(x_r,\tau;x_s) &= \int_{t1}^{t2}d(x_r,t-\tau;x_s)d^{obs}(x_r,t;x_s)dt \\
&= \int_{t1}^{t2}d(x_r,t;x_s)d^{obs}(x_r,t+\tau;x_s)dt,
\end{aligned}
\end{equation}
where $\tau$ is the time shift, $d(x_r,t;x_s)$ and $d^{obs}(x_r,t;x_s)$ are the synthetic and observed seismic data at for source $s$ and receiver $r$. The synthetic seismic data are dependent on the earth model parameter.

\subsection{Travel time difference/peak time shift}
The peak time shift is defined as the time shift that maximize the cross-correlation function
\begin{equation}
\delta \tau(x_r,x_s) =\underset{\tau} { \mbox{max} } ~C(x_r,\tau;x_s) = \underset{\tau} { \mbox{max} } \int_{t1}^{t2}d(x_r,t;x_s)d^{obs}(x_r,t+\tau;x_s)dt.
\end{equation}

\subsection{Implicit function between peak time shift and synthetic waveform data}
To obtain the connection between the peak time shift $\delta \tau(x_r,x_s)$ and model parameter $m(x)$, one needs to first estimate the relation between the peak time shift $\delta \tau(x_r,x_s)$ and synthetic data $d(x_r,t;x_s)$. The derivative of $C(x_r,\tau;x_s)$ with respect to $\tau$ is given by
\begin{equation}
\begin{aligned}
\dfrac{\partial C(x_r,\tau;x_s)}{\partial \tau} &= \int_{t1}^{t2}d(x_r,t;x_s)\dfrac{\partial d^{obs}(x_r,t+\tau;x_s)}{\partial \tau}dt \\
&= \int_{t1}^{t2}d(x_r,t;x_s)\dfrac{\partial d^{obs}(x_r,t+\tau;x_s)}{\partial (t+\tau)}\dfrac{\partial (t+\tau)}{\partial \tau}dt \\
& = \int_{t1}^{t2}d(x_r,t;x_s)\dot{d}^{obs}(x_r,t+ \tau;x_s) dt,
\end{aligned}
\end{equation}
where over-dot means time derivative. It is obvious that the derivative $\partial C(x_r,\tau;x_s)/\partial \tau$ vanishes at peak time shift $\tau = \delta \tau$.
\begin{equation}
\dfrac{\partial C(x_r,\tau;x_s)}{\partial \tau} \mid_{\tau = \delta \tau} =  \int_{t1}^{t2}d(x_r,t;x_s)\dot{d}^{obs}(x_r,t+ \delta\tau;x_s) dt = 0
\label{connection}
\end{equation}
Equation \ref{connection} indicates that the peak time shift $\delta \tau (x_r,x_s)$ is an implicit function of synthetic data $d(x_r,t;x_s)$. The connectivity can be denoted as
\begin{equation}
f[d(x_r,t;x_s),\delta \tau (x_r,x_s)] =  \int_{t1}^{t2}d(x_r,t;x_s)\dot{d}^{obs}(x_r,t+ \delta\tau;x_s) dt = 0.
\label{implicit}
\end{equation}

\subsection{Cross-correlation travel time misfit function for inversion}
The cross-correlation travel time misfit function for full-waveform inversion is defined as the sum of the square of the peak time shift
\begin{equation}
J = \dfrac{1}{2}\underset{s}{\sum}\underset{r}{\sum}\delta \tau(x_r,x_s)^2,
\end{equation}
where $\delta \tau(x_r,x_s)$ represents the peak time shift of observed and synthetic seismograms between source $s$ and receiver $r$. 

\subsection{Gradient of misfit function}

One of the key component of full-waveform inversion is the gradient of the misfit function
\begin{equation}
g(x) = \dfrac{\partial J}{\partial m(x)} = \underset{s}{\sum} \underset{r}{\sum} \dfrac{\partial \delta \tau(x_r,x_s)}{\partial m(x)}\delta \tau(x_r,x_s),
\label{gradient}
\end{equation}
the gradient is obtained by applying the Frechet derivative on the peak time shift. 

To obtain the Frechet derivative, we differentiate connectivity equation $f[d(x_r,t;x_s),\delta \tau(x_r,x_s)] = 0$ with respect to model parameter $m(x)$
\begin{equation}
\dfrac{\partial f [d(x_r,t;x_s),\delta \tau(x_r,x_s)]}{\partial d(x_r,t;x_s)}\dfrac{\partial d (x_r,t;x_s)}{\partial m (x)} + \dfrac{\partial f [d(x_r,t;x_s),\delta \tau(x_r,x_s)]}{\partial \delta \tau (x_r,x_s)}\dfrac{\partial \delta \tau (x_r,x_s)}{\partial m(x)} = 0
\end{equation}
The Frechet derivative is given by
\begin{equation}
\dfrac{\partial \delta \tau (x_r,x_s)}{\partial m (x)}= -\left[ \dfrac{\partial f (d,\delta \tau)}{\partial \delta \tau (x_r,x_s)}\right]^{-1}\left[ \dfrac{\partial f (d, \delta \tau)}{\partial d (x_r,t;x_s)}\dfrac{\partial d (x_r,t;x_s)}{\partial m (x)} \right]
\label{frechet}
\end{equation}
Using equation \ref{implicit}, we have
\begin{equation}
\begin{aligned}
\dfrac{\partial f (d,\delta \tau)}{\partial \delta \tau (x_r,x_s)} &= \int_{t1}^{t2}d(x_r,t;x_s)\dfrac{\partial \dot{d}^{obs}(x_r,t+ \delta \tau;x_s)}{\partial \delta \tau (x_r,x_s)}dt\\
& = \int_{t1}^{t2}d(x_r,t;x_s)\dfrac{\partial \dot{d}^{obs}(x_r,t+ \delta \tau;x_s)}{\partial (t+ \delta \tau)}\dfrac{\partial (t+ \delta \tau)}{\partial \delta \tau}dt\\
&=\int_{t1}^{t2}d(x_r,t;x_s)\ddot{d}^{obs}(x_r,t+\delta \tau;x_s)dt,
\end{aligned}
\end{equation}
and
\begin{equation}
\dfrac{\partial f(d,\delta \tau)}{\partial d(x_r,t;x_s)}\dfrac{\partial d(x_r,t;x_s)}{\partial m(x)} = \int_{t1}^{t2}\dfrac{\partial d(x_r,t;x_s)}{\partial m(x)}\dot{d}^{obs}(x_r,t+ \delta \tau;x_s)dt.
\end{equation}
With the above equations, the Frechet derivative can be expressed as \citep{luo1991}
\begin{equation}
\dfrac{\partial \delta \tau (x_r,x_s)}{\partial m(x)} = -\dfrac{\int_{t1}^{t2}\dfrac{\partial d(x_r,t;x_s)}{\partial m(x)}\dot{d}^{obs}(x_r,t+ \delta \tau;x_s)dt}{\int_{t1}^{t2}d(x_r,t;x_s)\ddot{d}^{obs}(x_r,t+\delta \tau;x_s)dt}.
\end{equation}

If assuming that $d(x_r,t;x_s)^{obs}$ and $d(x_r,t;x_s)$ are purely time shifted ($d^{obs}(x_r,t+\delta \tau;x_s)\approx d(x_r,t;x_s)$) \citep{marquering1999,dahlen2000,tromp2005,tape2007}, the Frechet derivative can be rewritten as
\begin{equation}
\dfrac{\partial \delta \tau(x_r,x_s)}{\partial m(x)} \approx -\dfrac{\int_{t1}^{t2}\dfrac{\partial d(x_r;t;x_s)}{\partial m(x)}\dot{d}(x_r,t;x_s)dt}{\int_{t1}^{t2}d(x_r,t;x_s)\ddot{d}(x_r,t;x_s)dt}.
\label{global}
\end{equation}
In this note, we follow the definition in equation \ref{global}. The gradient of misfit function (equation \ref{gradient}) changes to
\begin{equation}
\begin{aligned}
g(x) &= \dfrac{\partial J}{\partial m(x)} \\
&= \underset{s}{\sum} \underset{r}{\sum} \dfrac{\partial \delta \tau(x_r,x_s)}{\partial m(x)}\delta \tau(x_r,x_s) \\
&\approx -\underset{s}{\sum} \underset{r}{\sum} \delta \tau(x_r,x_s) \dfrac{\int_{t1}^{t2}\dfrac{\partial d(x_r,t;x_s)}{\partial m(x)}\dot{d}(x_r,t;x_s)dt}{\int_{t1}^{t2}d(x_r,t;x_s)\ddot{d}(x_r,t;x_s)dt}\\
&= -\underset{s}{\sum} \underset{r}{\sum} \int_{t1}^{t2} \dfrac{\partial d(x_r,t;x_s)}{\partial m(x)}\left[\dot{d}(x_r,t;x_s)\dfrac{\delta \tau(x_r,x_s)}{N}\right]dt
\end{aligned}
\label{gradient_cc}
\end{equation}
where $N = \int_{t1}^{t2}d(x_r,t;x_s)\ddot{d}(x_r,t;x_s)dt$ is normalizer and $\partial d(x_r,t;x_s)/\partial m(x)$ is the Frechet derivative of waveform data $d(x_r,t;x_s)$ with respect to $m(x)$.
This formulation is very similar to the gradient of the $\ell_2$ norm waveform misfit function
\begin{equation}
g_{\ell_2}(x) = -\underset{s}{\sum} \underset{r}{\sum} \int_{t1}^{t2} \dfrac{\partial d(x_r,t;x_s)}{\partial m(x)}\left[\delta d(x_r,t;x_s)\right]dt
\label{gradient_ls}
\end{equation}
where $\delta d(x_r,t;x_s) = d^{obs}(x_r,t;x_s) - d(x_r,t;x_s)$ is the waveform difference. The difference between equation \ref{gradient_ls} and equation \ref{gradient_cc} is the replacement of $[\delta d(x_r,t;x_s)]$ by $[\dot{d}(x_r,t;x_s)\dfrac{\delta \tau(x_r,x_s)}{N}]$.

\subsection{Adjoint source}
Instead of explicitly construct the Frechet derivative, the adjoint-state method is usually used to compute the action of Frechet derivative on data residuals. 
We first look at the Frechet derivative $\partial d(x_r,t;x_s)/\partial m(x)$. 

The elastic wave equation can be abstractly written as
\begin{equation}
{\bf S}{\bf u} = {\bf f},
\label{wave}
\end{equation}
where ${\bf S}$ denotes wave equation operator, ${\bf u}$ is wavefield vector and ${\bf f}$ is source vector. 
The solution of the wave equation is 
\begin{equation}
{\bf u} = {\bf S}^{-1}{\bf f}.
\end{equation}
The corresponding Green's function of wave equation \ref{wave} follows  
\begin{equation}
S(x,t)G(x,t;x',t') = \delta (x-x') \delta(t-t')
\end{equation}
The integration representation of the solution of the wave equation is
\begin{equation}
u(x,t) = \int_{0}^t\int_V G(x,t;x',t')f(x',t')d^3x'dt'
\end{equation}

Assuming that the Green's function is time invariant 
\begin{equation}
G(x,t;x',t') = G(x, t-t';x',0) =G(x, 0;x',t'-t) 
\end{equation}
The wavefield can be written as
\begin{equation}
\begin{aligned}
u(x,t) &= \int_{0}^t\int_V G(x, t-t';x',0)f(x',t')d^3x'dt'\\
&=\int_V G(x,t;x',0) * f(x',t)d^3x'
\end{aligned}
\end{equation}
The wavefield due to a point source located at $x'=x_s$ is
\begin{equation}
u(x,t;x_s) = G(x,t;x_s,0) * f(t;x_s),
\end{equation}
where $f(x',t') = \delta(x'-x_s)f(t;x_s)$.

The data recorded at receivers are a sampled version of the wavefield
\begin{equation}
d(x_r,t;x_s) = u(x_r,t;x_s) =G(x_r,t;x_s,0) * f(t;x_s)
\end{equation}

Differentiate the wave equation with respect to model vector
\begin{equation}
\dfrac{\partial {\bf S}}{\partial {\bf m}} {\bf u} + {\bf S}\dfrac{\partial {\bf u}}{\partial {\bf m}} = {\bf 0},
\end{equation}
where the right side of the equation vanishes because the source term is not dependent on the model parameter.
The Frechet derivative of ${\bf u}$ with respect to ${\bf m}$ can be expressed as
\begin{equation}
\dfrac{\partial {\bf u}}{\partial {\bf m}} = -{\bf S}^{-1}\left(\dfrac{\partial {\bf S}}{\partial {\bf m}}{\bf u} \right)
\end{equation}
The Born scattered wavefield $\delta {\bf u}$ due to scatter $\delta {\bf m}$ can be written as
\begin{equation}
\delta {\bf u} = \dfrac{\partial {\bf u}}{\partial {\bf m}}\delta {\bf m} = -{\bf S}^{-1}\left(\dfrac{\partial {\bf S}}{\partial {\bf m}}{\bf u} \right)\delta {\bf m},
\label{born}
\end{equation}
where ${\bf u}$ is the incident wavefield. Using the Green's function, equation \ref{born} changes to
\begin{equation}
\delta u(x,t) = -\int_VG(x,t;x',0)*\left[\dfrac{\partial S(x',t)}{\partial m(x')} u(x',t)\delta m(x')\right]d^3x'
\end{equation}
The scattered data recorded at the location of receivers due to incident wavefield excited by source $s$ are
\begin{equation}
\delta d(x_r,t;x_s) = \delta u(x_r,t;x_s) = -\int_VG(x_r,t;x,0)*\left[\dfrac{\partial S(x,t)}{\partial m(x)} u(x,t;x_s)\delta m(x)\right]d^3x
\end{equation}
The integration kernel of Frechet derivative $\partial d/\partial m$ is
\begin{equation}
\dfrac{\partial d(x_r,t;x_s)}{\partial m(x)} = -G(x_r,t;x,0) * \left[\dfrac{\partial S(x,t)}{\partial m(x)}u(x,t;x_s) \right]
\end{equation}

The gradient of cross-correlation travel time misfit function equation \ref{gradient_cc} can be written as
\begin{equation}
\begin{aligned}
 g(x) &= -\underset{s}{\sum} \underset{r}{\sum} \int_{t1}^{t2} \dfrac{\partial d(x_r,t;x_s)}{\partial m(x)}\left[\dot{d}(x_r,t;x_s)\dfrac{\delta \tau(x_r,x_s)}{N}\right]dt\\
 &=\underset{s}{\sum} \underset{r}{\sum} \int_{t1}^{t2}\left\{G(x_r,t;x,0) * \left[\dfrac{\partial S(x,t)}{\partial m(x)}u(x,t;x_s) \right] \right\} \left[\dot{d}(x_r,t;x_s)\dfrac{\delta \tau(x_r,x_s)}{N}\right]dt.
\end{aligned}
  \label{gradient_green}
 \end{equation}
Using the property of convolution
\begin{equation}
\int[f(t)*g(t)]h(t)dt = \int g(t)[f(-t)*h(t)]dt,
\end{equation}
equation \ref{gradient_green} can be changed to
\begin{equation}
\begin{aligned}
g(x) &= \underset{s}{\sum} \underset{r}{\sum} \int_{t1}^{t2}\left[\dfrac{\partial S(x,t)}{\partial m(x)}u(x,t;x_s) \right] \left\{G(x_r,-t;x,0) * \left[\dot{d}(x_r,t;x_s)\dfrac{\delta \tau(x_r,x_s)}{N}\right]  \right\}dt\\
&= \underset{s}{\sum} \underset{r}{\sum} \int_{t1}^{t2}\left[\dfrac{\partial S(x,t)}{\partial m(x)}u(x,t;x_s) \right] \left\{G(x,0;x_r,t) * \left[\dot{d}(x_r,t;x_s)\dfrac{\delta \tau(x_r,x_s)}{N}\right]  \right\}dt,
\end{aligned}
\end{equation}
where the time invariant property ($G(x_r,-t;x,0)= G(x_r,0;x,t)$) and reciprocity of Green's function is used ($G(x_r,0;x,t) = G(x,0;x_r,t)$).
The physical meaning of the term in the curly bracket is taking the adjoint source, placing it in the receiver position $x_r$ and propagating it in reversing time direction ($t \rightarrow 0$).
The adjoint source is given by the formulation
\begin{equation}
s_{adj}= \dot{d}(x_r,t;x_s)\dfrac{\delta \tau(x_r,x_s)}{N}.
\end{equation}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%EXAMPLES%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}
\subsection{Synthetic example}
In the synthetic data example, the observed data is a Ricker wavelet with peak frequency $1$ Hz. The synthetic data is a shifted and amplitude reduced version of the observed data. The time shift measured by the cross-correlation is $-0.2$ s. It means that the observed data is $0.2$ s advance than the synthetic data. The value of the cross-correlation travel time misfit function is $0.02$. 

\subsection{Real data example}
The observed seismogram is from earthquake in Berkeley January 4, 2018. The recording station is WENL - Wente Vineyards, Livermore, CA. The observed data is low-pass filtered with corner frequency $0.02$ Hz. The synthetic data is simulated using $SW4$. 
The time shift measure by cross-correlation is $1.025$ s. The value of the misfit function is $0.525$


\renewcommand{\figdir}{Fig} % figure directory
\multiplot{1}{Figure1,Figure2}{width=0.8\textwidth}{(a) The observed and synthetic data. The observed data is $0.1$ sec in advance of the synthetic data. (b) The normalized adjoint source for cross-correlation travel time misfit.}

\plot{Figure3}{width=\textwidth}{The Jan. 4, 2018 Berkeley earthquake data recorded at station WENL.The observed (blue) and synthetic (red) data. The data are low-pass filtered with corner frequency $0.02$ Hz.}

\multiplot{1}{Figure4,Figure5,Figure6,Figure7}{width=0.8\textwidth}{(a) Horizontal component of the synthetic and real earthquake data (Jan 4, 2018 Berkeley event). (b) Windowed version of the data. (c) The normalized adjoint source for cross-correlation travel time misfit. (d) The observed data is shifted by the time delay estimated from cross-correlation for comparison.}



\bibliographystyle{seg}  % style file is seg.bst
\bibliography{cc}
\end{document}